Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-07-12 manuscript/response_to_reviews.Rmd

L287-299: it would be very useful to use scatterplots to show the effects of traits on Rt in a figure.

[*issue #71*](https://github.com/SCBI-ForestGEO/McGregor_climate-sensitivity-variation/issues/71)

L291-294: but this is problematic considering that you only had two diffuse-porous species.

**We have made changes to the analysis that resulted in dropping ring porosity from the full models. While we agree that it is problematic that we only have two ring-porous species, we believe that it's worth including because this engages with, and rejects, the existing understanding that in this biome, ring-porous species are more drought tolerant.**

@@

L305-306: not for the 1977 and 1999 droughts, according to Table 5.

L314-316: or simply that your sample size was rather low. See also main comments above.

L329-331: do you have any information on the demographic effects (e.g., mortality) of the three droughts studied?

**Unfortunately, no. All we could say is that the 1999 drought didn't result in a big loss of live biomass.**

@@

Figure 1: please explain all acronyms.

**We now specify that these codes are given in Table 3.**

Figure 2: where are the leaf hydraulic traits shown (cf. caption)?

**It was an error that this was mentioned in the caption; now removed.**

@@

Table S1: why only 11 species? Some of the fits are really low…

**Bark thickness measurements were drawn from an existing data set, and we did not make additional measurements because the trait is unlikely to have meaningful influence on results. This study already goes above and beyond the existing norm for accounting for bark thickness in DBH reconstructions; typically there is no correction for changes in bark thickness as the tree ages.**

### 2020-07-12 manuscript/climate_sensitivity_SI.Rmd
worked on reviewer responses (misc small changes)

\newpage
![Figure S3: Height by canopy position across the three focal droughts and in the year of measurement (2018)](tables_figures/publication/figureS3_height_plot_analysis.png){width=600px}

>>>>

\newpage
![Figure S3: Height (from reconstructed DBH) by canopy position across the three focal droughts and in the year of measurement (2018)](tables_figures/publication/figureS3_height_plot_analysis.png){width=600px}

### 2020-07-12 manuscript/climate_sensitivity_draft.Rmd
worked on reviewer responses (misc small changes)

Hydraulic traits were collected in August 2018 (Table 3). We sampled small sun-exposed branches up to eight meters above ground from three individuals of each species in and around the ForestGEO plot. Sampled branches were re-cut under water at least two nodes above the original cut and re-hydrated overnight in covered buckets under opaque plastic bags before measurements were taken. Rehydrated leaves taken towards the apical end of the branch (n=3 per individual: small, medium, and large) were scanned, weighed, dried at 60$^\circ$ C for $\ge$ 48 hours, and then re-scanned and weighed. Leaf area was calculated from scanned images using the LeafArea R package [@R-LeafArea]. $LMA$ was calculated as the ratio of leaf dry mass to fresh area. $PLA_{dry}$ was calculated as the percent loss of area between fresh and dry leaves. wood density was calculated for ~1cm diameter stem samples (bark and pith removed) as the ratio of dry weight to volume, which was estimated using Archimedes’ displacement. We used the rapid determination method of @bartlett_rapid_2012 to estimate osmotic potential at turgor loss point ($\pi_{tlp}$). Briefly, two 4 mm diameter leaf discs were cut from each leaf, tightly wrapped in foil, submerged in liquid nitrogen, perforated 10-15 times with a dissection needle, and then measured using a vapour pressure osmometer (VAPRO 5520, Wescor, Logan, UT, USA). Osmotic potential ($\pi_{osm}$) given by the osmometer was used to estimate ($\pi_{tlp}$) using the equation $\pi_{tlp}=0.832 \pi_{osm} ^{-0.631}$ [@bartlett_rapid_2012].

@@

The leaf hydraulic traits $PLA_{dry}$ and $\pi_{tlp}$ were linked to drought responses, whereas the other traits considered had insignficant and/or inconsistent correlations to $Rt$ (Tables 1,4,5). In the single-variable tests, $LMA$ and wood density were never significantly associated with $Rt$ (Table 4) and were excluded from the full models. In contrast, xylem porosity, $PLA_{dry}$, and $\pi_{tlp}$ all explained modest amounts of variation ($\Delta$AICc > 1.0) during at least one of the three droughts (Table 4). Xylem porosity was not significant for all droughts combined and had contrasting effects in the individual droughts: whereas ring-porous species had higher $Rt$ than diffuse- and semi-ring- porous species in the 1966 and 1999 droughts, they had lower $Rt$ in 1977 (Tables 4,5). $PLA_{dry}$ was a strong predictor for 1966 and all droughts combined, with consistently negative coefficients (Table 4). Similarly, $PLA_{dry}$ was consistently included, with negative coefficient, in full models for the three droughts combined and for the 1966 and 1977 droughts individually (Table 5). $\pi_{tlp}$ was not significant in any single-variable tests; however, coefficients were consistently negative (Table 4) and $\pi_{tlp}$ was included in the top full model for all droughts combined and for the 1977 and 1999 droughts individually (Table 5).

@@

Figure S3: Height by canopy position across the three focal droughts and in the year of
measurement (2018)

>>>>

Hydraulic traits were collected in August 2018 (Table 3). We sampled small sun-exposed branches up to eight meters above ground from three individuals of each species in and around the ForestGEO plot. Sampled branches were re-cut under water at least two nodes above the original cut and re-hydrated overnight in covered buckets under opaque plastic bags before measurements were taken. Rehydrated leaves taken towards the apical end of the branch (n=3 per individual: small, medium, and large) were scanned, weighed, dried at 60$^\circ$ C for $\ge$ 48 hours, and then re-scanned and weighed. Leaf area was calculated from scanned images using the LeafArea R package [@R-LeafArea]. $LMA$ was calculated as the ratio of leaf dry mass to fresh area. $PLA_{dry}$ was calculated as the percent loss of area between fresh and dry leaves. Wood density was calculated for ~1cm diameter stem samples (bark and pith removed) as the ratio of dry weight to fresh volume, which was estimated using Archimedes’ displacement. We used the rapid determination method of @bartlett_rapid_2012 to estimate osmotic potential at turgor loss point ($\pi_{tlp}$). Briefly, two 4 mm diameter leaf discs were cut from each leaf, tightly wrapped in foil, submerged in liquid nitrogen, perforated 10-15 times with a dissection needle, and then measured using a vapour pressure osmometer (VAPRO 5520, Wescor, Logan, UT, USA). Osmotic potential ($\pi_{osm}$) given by the osmometer was used to estimate ($\pi_{tlp}$) using the equation $\pi_{tlp}=0.832 \pi_{osm} ^{-0.631}$ [@bartlett_rapid_2012].

@@

Wood density, $LMA$, and xylem porosity were all poor predictors of $Rt$ (Tables 1,4,5). 
Wood density and $LMA$ were never significantly associated with $Rt$ in the single-variable tests (Table 4) and were therefore excluded from the full models. 
Xylem porosity was also excluded from the full models, as it had no significant influence for all droughts combined and had contrasting effects in the individual droughts: whereas ring-porous species had higher $Rt$ than diffuse- and semi-ring- porous species in the 1966 and 1999 droughts, they had lower $Rt$ in 1977 (Table **4**). 
It is noteworthy that the two diffuse-porous species in our study, *Liriodendron tulipifera* and *Fagus grandifolia*, were at opposite ends of the $Rt$ spectrum (Fig. **2**), further refuting the idea that xylem porosity is a useful predictor of $Rt$.

In contrast, $PLA_{dry}$, and $\pi_{tlp}$ were linked to drought responses (Tables 1,4,5). 
Both had consistent signs across all droughts and explained modest amounts of variation ($\Delta$AICc > 1.0) during at least one of the three droughts (Table 4), qualifying them as candidate variables for the full model. 
$PLA_{dry}$ was consistently included, with negative coefficient, in full models for the three droughts combined and for the 1966 and 1977 droughts individually (Table 5). 
$\pi_{tlp}$ was included in the top full model for all droughts combined and for the 1977 and 1999 droughts individually (Table 5).

@@

Figure S3: Height (from reconstructed DBH) by canopy position across the three focal droughts and in the year of
measurement (2018)

### 2020-07-12 manuscript/tables&figures.Rmd
worked on reviewer responses (misc small changes)

![**Figure 3. Height profiles in growing season climatic conditions, tree heights by crown position, and leaf hydraulic traits** The top row shows averages ($\pm$ SD) of daily maxima and minima of (a) wind speed, (b) relative humidity ($RH$), and (c) air temperature ($T_{air}$) averaged over each month of the peak growing season (May-August) from 2016-2018. In these plots, heights are slightly offset for visualization purposes. Also shown is (d) 2018 tree heights by canopy position (see Table 2 for codes). In all plots, the dashed horizontal line indicates the 95th percentile of tree heigts in the ForestGEO plot.](tables_figures/publication/Figure3_NEON_vars_height_profile.png){width=500px}

>>>>

![**Figure 3. Height profiles in growing season climatic conditions and tree heights by crown position** The top row shows averages ($\pm$ SD) of daily maxima and minima of (a) wind speed, (b) relative humidity ($RH$), and (c) air temperature ($T_{air}$) averaged over each month of the peak growing season (May-August) from 2016-2018. In these plots, heights are slightly offset for visualization purposes. Also shown is (d) 2018 tree heights by canopy position (see Table 2 for codes). In all plots, the dashed horizontal line indicates the 95th percentile of tree heigts in the ForestGEO plot.](tables_figures/publication/Figure3_NEON_vars_height_profile.png){width=500px}
